\section{PROGRAM ACTIVITIES}
\label{sec:activities}

\subsection{PI and Co-PIs Involvement}

\bits
\item First year. PI: program coordination and curriculum planning (within the department and between the other departments participating in ACMS). Design in collaboration with the ACMS leadership and the departments involved of the set of courses to form the new curriculum. Ensure consistency of the prerequisites between courses in the track, as well as feasible paths to timely graduation exist. Plan and start on \statcl~ course notes. Co-PI's: Development of a plan for the core Python numerical and data structures libraries to be taught.
Beginning of the development of course notes for \statcl.
\item Second year, PI, Co-PIs: Evaluation of the first phase and
  incorporation of the lessons learned. Development of the bulk of the
  STAT 391 course notes. Supervision of the undergraduate research
  seminar.  Selecting/curating (with RAs) of the data sets to be used
  in \statcl. Organization of first workshop.
\item Third year: Evaluation of second phase, fine-tuning of the
  curriculum and program requirements. Exploration of the
  possibilities to open this pathway to math majors. Organization of
  second workshop.  \eits

\subsection{Undergraduate involvement}
\label{sec:ugradseminar}
%The PI and all co-PI's have extensive track records of involving
%undergrads in research. Undergraduates who take either of the courses will be involved in
%research projects either (1) along with the funded graduate students,
%supervised by the PI, co-PI's and the graduate students or (2) in the
%UW units that provide data sets.
All undergraduates involved in research under this project, along with
their supervising graduate students, will participate in an {\em
  Undergraduate \cdse\ Research Seminar} where they will present and
discuss their work.  The seminar will be open to any other
undergraduate students involved in research with Statistics faculty
who would like to participate, as well as any other undergraduates
across campus interested in \cdse~ research. The seminar's goal is to
be a forum where research experiences are shared, research teams are
formed, new research projects are started, and familiarity with the
strategic and social aspects of research is gained.
In addition to technical presentations by undergraduate
students, the seminar will include:
\bits
\item meetings where faculty (from sciences for example) present quantitative data analysis research projects to recruit interested undergraduates
\item  mentorship sessions by guest speakers, UW graduate students and faculty, on career options in research, applying to graduate school, how to work in a team, how to approach a potential research advisor, how to approach a new research project
\item presentations from outside speakers (e.g local businesses such as Amazon,
  Microsoft, and Google) on career opportunties in \cdse. As much as possible, we will focus on drawing a diverse group of guest speakers, including women and underrepresented minorities.
\eits
Some of the seminar's activities and goals overlap with the goals of {\em Statistics and Probability Association}\footnote{{\tt http://students.washington.edu/spassc/index.html}}, and the Pre-MAP program\footnote{\tt http://www.astro.washington.edu/users/premap/}, 
and we plan to do these jointly, reinforcing the  benefits for all groups involved.

\subsection{Graduate students involvement}
\label{sec:activities-grad}

For the statistics graduate students funded by this grant, we envision 
\bits
\item to fund several students for a relatively short time (2 quarters
  to 1 year). We adopt this ``rotation'' plan recognizing that developing 
tools cannot be the only part of a PhD work in statistics. 
This ``rotation'' model will also assure a truly functional data
infrastructure, as each departing graduate student 
will have to ensure smooth transition to her/his successor.)
\item Students will help develop the software and data infrastructure for
  the program, search for available data sets, curate them, write
  preprocessing software if necessary, and design tasks and exercises. 
  At the same time, students will get training for analyzing large data sets. 
\item Students will be advised by the PI, by another Statistics faculty, or by another UW scientist with interests in statistical analysis of big data. Gradually, a research problem will be formulated and the student will
focus on the data and methodology relevant to the chosen problem. After this ``apprenticeship'', 
some students will continue their research supervised by other advisors. 
\item The students will also be strongly recommended to TA the STAT 391 class, thus rounding their preparation and providing self-confidence. 
\item This program will enable statistics PhD students to play useful roles in other NSF funded initiatives at UW. For 
%instance, in the CSNE, with with the PI is involved, where the data collected, far from reaching Tera byte sizes, represents a %daunting challenge for the average Statistics graduate student who hasn't accquired a CS degree before. T
example, this plan harmonizes with the IGERT plan of offering graduate students the experience of working with domain scientists, and with outside big data companies via internships. 
\item Finally, the graduate students will participate in the Undergraduate Research Seminar, and will mentor 1-2 undergraduates. 
 \eits

\subsection{Evaluation plan}

The evaluation of the outcomes of this proposal will be performed by
the UW Office of Educational Assessment (OEA), with respect to the
program goals listed in Section \ref{sec:key-aims}.

The evaluation plan includes the use of both qualitative and
quantitative methods for the purposes of improving offerings across
years of the grant (formative evaluation) and assessing the extent to
which the grant-funded activities are effective in meeting projected
outcomes (summative evaluation).  To assess student learning, OEA will
consult with instructors of the proposed courses to ensure that
in-course assessments (e.g., homework assignments, exams, capstone
projects) align with the course learning goals.  


\underline{Year 1} OEA will work with the project investigators to
develop a full logic model for the program. At this point, it is
anticipated that the summative evaluation will focus on the immediate
and long-term outcomes (Section \ref{sec:key-aims}.  In addition, OEA
will gather basic information about participation in activities
(enrollment in new courses, and new major), and demographics where
relevant to building a diverse workforce.

Other evaluation activities: design of in-course assessments, pre- and
post-surveys, and end-of-course focus groups with students enrolled in
new courses, interviews with students involved in research, creation
of tracking database for all student participants.

\underline{Years 2 and 3} activities: pre- and post-surveys for
students in \statcl, \astrocl, interviews with students involved in
research, follow-up surveys, survey for faculty taking part in
workshop.  

Deliverables, each year: reports on all student data, reports on
faculty workshops, two evaluation meetings with PIs for formative
review of evaluation results.

