\documentstyle[11pt,aaspp4]{article}

%% latex ZIcv;  dvips -o ZIcv.ps ZIcv; ps2pdf ZIcv.ps    

\begin{document}

\def\ref{\hangindent 2pc \hangafter 1}
\def\Bline#1{\hskip 0.3in $\bullet$ #1 \\ \smallskip}
\def\Sline#1{\hskip 0.6in #1 \\}
\def\line#1{#1 \\}


{\bf  Biographical Sketch for \v{Z}eljko Ivezi\'{c}} 
\smallskip


\begin{tabbing}
% Set tabs:
1234567 \= 12345678901234567890123456  \= 12345678901234567  \= 1234567890 \= \kill
\hskip 0.4in {\bf  Professional Preparation} \smallskip
\\
\>  University of Zagreb   \>  Mechanical Eng.  \> B.Sc.   \> 1990 \\
\>  University of Zagreb   \>  Physics          \> B.Sc.   \> 1991 \\
\>  University of Kentucky \>  Physics          \> Ph.D.   \> 1995 \\
\>  University of Kentucky \>  Astrophysics     \> Postdoc \> 1995-1996 \\
\>  Princeton University   \>  Astronomy        \> Postdoc \> 1997-1999 \\
\end{tabbing}

\vskip -0.8in
\phantom{x}
\vskip -0.8in

\begin{tabbing}
% Set tabs:
1234567 \= 12345678901234567890123456  \= 12345678901234567890123456 \= \kill
\hskip 0.4in {\bf  Appointments} \smallskip
\\
\> University of Washington  \> Professor  \>  2010-present \\ 
\> University of Washington  \> Associate Professor  \>  2006-2010\\ 
\> University of Washington  \> Assistant Professor  \>  2004-2006 \\ 
\> Princeton University   \> H.N. Russell Fellow     \>  2002-2004 \\ 
\> City University of New York   \> Assistant Professor   \>  1998-2002 \\ 
\> Princeton University   \> Research Staff Member   \>  1999-2002 \\ 
\> Princeton University   \> Postdoctoral Researcher \>  1997-1999 \\
\> University of Kentucky \> Postdoctoral Scholar    \>  1995-1996 \\
\> University of Kentucky \> Research Assistant      \>  1991-1995 \\
\> University of Zagreb   \> Teaching Assistant      \>  1990-1991 \\
\end{tabbing}

\vskip -0.2in

\hskip 0.1in {\bf Publications} 

\Bline{ Related Publications\hfil}
\vskip -0.3in

\ref 
Kimball, A., Ivezi\'{c}, \v{Z}., Wiita \& Schneider 2011,  {\it Correlations of Quasar Optical Spectra with Radio 
Morphology}, Astronomical Journal, 141, 182.

\ref 
MacLeod, C., Ivezi\'{c}, \v{Z}., et al. 2010,  {\it Modeling the Time Variability of SDSS Stripe 82 Quasars as Damped
Random Walk}, Astrophysical Journal, 721, 1014.

\ref 
McGurk, R., Kimball, A. \& Ivezi\'{c}, \v{Z}. 2010,  {\it Principal Component Analysis of SDSS Stellar Spectra}, 
Astronomical Journal, 139, 1261.

\ref 
Palaversa, L.,  Ivezi\'{c}, \v{Z}.,  et al. 2013,   {\it Exploring the Variable Sky with LINEAR. III. Classification of 
   Periodic Light Curves}, Astronomical Journal, 146, 101.

\ref 
Parker, A., Ivezi\'{c}, \v{Z}., et al. 2008,  {\it The Size Distributions of Asteroid Families in the SDSS Moving 
Object Catalog 4}, Icarus, 198, 138.

\newpage
\Bline{ Other Significant Publications\hfil}
\vskip -0.3in

\ref
Ivezi\'{c}, \v{Z}., Beers, T.C. \& Juri\'{c}, M. 2012, {\it Galactic Stellar Populations in the Era 
of the Sloan Digital Sky Survey and Other Large Surveys}, Annu. Rev. Astron. Astrophys., 50, 251.

\ref
Ivezi\'{c}, \v{Z}., Tyson, J.A., Acosta, E., et al. 2008, {\it LSST: from Science Drivers to Reference 
Design and Anticipated Data Products}, arXiv:0805.2366.

\ref
Nenkova, M., Ivezi\' c, \v Z., \& Elitzur, M. 2002, {\it Dust Emission from 
Active Galactic Nuclei}, Astrophysical Journal, 570, L9.

\ref 
Ivezi\' c, \v Z., Menou, K., Knapp, G.R. et al. 2002, {\it Optical and Radio 
Properties of Extragalactic Sources Observed by the FIRST Survey and
the Sloan Digital Sky Survey}, Astronomical Journal, 124, 2364.

\ref
Ivezi\' c, \v Z., Tabachnik, S., Rafikov, R., et al. 2001, {\it Solar System 
Objects Observed in the Sloan Digital Sky Survey Commissioning Data},
Astronomical Journal, 122, 2749.


\hskip 0.1in {\bf Synergistic Activities}

\ref
LSST Project Scientist and the chair of LSST Project Science Team.

\ref
Co-developer of the code DUSTY for solving the radiative transfer problem 
for a source embedded in a dusty envelope. The code is publicly available 
(http://www.pa.uky.edu/$\sim$moshe/dusty) and is used world-wide by over 
50 research groups.

\ref
Co-developer of processing algorithms for astronomical image analysis,
and of a test suite for quality assurance, in support of the Sloan Digital Sky Survey.

\ref
Co-author of the publicly available Sloan Digital Sky Survey Moving 
Object Catalog (http://www.sdss.org/science/index.html) (the largest 
catalog of multi-color asteroid photometry). 


\hskip 0.1in {\bf Collaborators \& Other Affiliations}

\Bline{Collaborators} 
\ref
\hskip -0.08in The SDSS Collaboration, The LSST Collaboration (the full list 
 is much longer than 2 pages). The closest collaborators include Moshe Elitzur,
Jill Knapp, Jim Gunn, Robert Lupton, Andrew Becker, Gordon
Richards, Pat Hall, Daniel Vanden Berk, Xiaohui Fan, and Paul Wiita.

\Bline{Graduate and Postdoctoral Advisors}
\ref
\hskip -0.08in Moshe Elitzur, University of Kentucky, Gillian Knapp, Princeton University

\Bline{Thesis Advisor and Postgraduate-Scholar Sponsor: Amy Kimball,
Branimir Sesar, }
\ref
\hskip -0.08in Michael Solontoi, Chelsea MacLeod, Sarah Loebman, Lynne Jones, Peter Yoachim.

\end{document}
